Large Matrix Manipulation for Unfolding Optimization

### 2019-05-03 NOvA/numu/unfolding_optimization/large_matrix_manipulation.tex
Adding collinearity study.
Adding a less technical version.

\begin{frame}{Overview}
  \begin{itemize}
    \item Currently, the number of iterations we use for unfolding was optimized in 1D neutrino energy only, but not in the full 3D $(\cos\theta_\mu,T_\mu,E_\nu)$ space.
    \item I have been trying to optimize the number of iterations for iterative unfolding in the full 3D space.
    \item In this study, RooUnfold is used to calculate the covariance matrix between bins.
    \item The total number of bins in $(\cos\theta_\mu,T_\mu,E_\nu)$ is $14\times 22\times 22=6776$. As a result, the covariance matrix is of size $6776\times 6776$.
    \item It turns out large matrices are hard to manipulate numerically. 
  \end{itemize}
  \bigskip
  \centering
  \textcolor{red}{Welcome to numerical linear algebra!}
\end{frame}

@@

\begin{frame}{Removing Zero Rows and Columns Does Not Help!}
  \begin{center}
    \includegraphics[height=.6\textheight]{figures/iter2tol1e-15.png}
  \end{center}
  \begin{itemize}
    \scriptsize
    \item Plots are with 2 iterations. Similar situation applies to all iterations.
    \item The reason is a large matrix becomes nearly singular if it contains highly correlated rows or columns. 
    \item The adjacent columns of a covariance matrix are very often highly correlated.
    \item Of course, smaller matrices are always better. Will move to using the zero suppressed matrices.
  \end{itemize}
\end{frame}

>>>>

\begin{frame}{Overview}
  \begin{itemize}
    \item I have been dealing with covariance matrices of size $6776\times 6776$ for optimizing the number of iterations in iterative unfolding.
    \item In this study, RooUnfold is used to calculate the covariance matrix.
    \item Matrix manipulation is done with \texttt{PyROOT}, \texttt{root\_numpy}, and \texttt{numpy.linalg}.
    \item It turns out large matrices are hard to manipulate numerically.
  \end{itemize}
  \bigskip
  \centering
  \textcolor{red}{Welcome to numerical linear algebra!}
\end{frame}

@@

\begin{frame}{Removing Zero Rows and Columns Does Not Help!}
%  \begin{center}
%    \includegraphics[height=.6\textheight]{figures/iter2tol1e-15.png}
%  \end{center}
  \begin{columns}
    \begin{column}{.68\textwidth}
      \includegraphics[width=\textwidth]{figures/iter2tol1e-15.png}
    \end{column}
    \begin{column}{.32\textwidth}
      \includegraphics[width=\textwidth]{figures/collinearity.pdf}
    \end{column}
  \end{columns}
  \begin{itemize}
    \scriptsize
    \item Plots are with 2 iterations. Similar situation applies to all iterations.
    \item The reason is a large matrix becomes nearly singular if it contains highly correlated rows or columns. 
    \item The adjacent columns of a covariance matrix are very often highly correlated.
    \item Of course, smaller matrices are always better. Will move to using the zero suppressed matrices.
  \end{itemize}
\end{frame}
